---
title: "02 - Other Trivia Plots"
author: "Alex Pinard"
date: "2020-08-18"
output: html_document
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE)
```

First, load the trivia_stats.R source file, which contains all functions for generating statistics:

```{r, results="hide", message=FALSE}
source("trivia_stats.R")
```

## Player Similarity Table

This chart measures how similar players scores are on a given round (after adjusting for round difficulty). Basically - bias chart measures how well a player does on another player's rounds, similarity table measures how similarly they perform on all the rest of the rounds.

```{r}
player_similarity_table()
```

This chart is a 3D representation of the similarity table - obtained using Bayesian Principal Components Analysis (BPCA), using the 2nd-4th principal components (the first is assumed to largely deal with round difficulty, not with player interactions).

```{r}
plot_3d_similarity_plotly()
```

## Bias Heatmap

Each column represents the scores on one person's round that week. Higher scores in a cell mean that the player tends to gain more points than other players on that creator's rounds.

Estimated Score Per Round = Average Score + Round Difficulty + Bias

```{r}
bias_table_heatmap()

```

## Scores on a Creator's Round

This plot displays the average score on each round by week. Hover over a point to see the creator for that round.

```{r}
creator_overlay_plot()
```

Each of the following plots shows the scores on the rounds created by one person each week. The red line represents the average for that week, and the blue dots are individual scores that week (they need to be labeled at some point):

```{r paged.print=TRUE}
generate_creator_plot("Alex")
generate_creator_plot("Chris")
generate_creator_plot("Dad")
generate_creator_plot("Drew")
generate_creator_plot("Ichigo")
generate_creator_plot("Jeff")
generate_creator_plot("Jenny")
generate_creator_plot("Megan")
generate_creator_plot("Mom")
generate_creator_plot("Zach")
```

## Scores by Player

Each of the following plots shows the final score for a single player each week. Notice the p-value in the bottom right side of each output summary. If this value is less than 0.05, this result is considered unlikely to happen by pure chance and may indicate an underlying trend. 

```{r paged.print=TRUE}
generate_player_plot("Alex")
generate_player_plot("Chris")
generate_player_plot("Dad")
generate_player_plot("Drew")
generate_player_plot("Ichigo")
generate_player_plot("Jeff")
generate_player_plot("Jenny")
generate_player_plot("Megan")
generate_player_plot("Mom")
generate_player_plot("Zach")
```