---
title: "01 - Trivia Plots"
author: "Alex Pinard"
date: "2020-08-18"
output: html_document
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE)
```

First, load the trivia_stats.R source file, which contains all functions for generating statistics:

```{r, results="hide", message=FALSE}
source("trivia_stats.R")
```

## Joker Table

The only practical chart I have made so far. This chart takes into account both the creators's average round difficulty and your own bias toward or against that creator to given you an overall sense of how much better than average you tend to do on rounds made by that creator (and therefore, who you should pick to joker today).

Estimated Score Per Round = Average Score + Round Difficulty + Bias


```{r}
joker_table_heatmap()
scores_boxplot_player("Alex")
scores_boxplot_player("Chris")
scores_boxplot_player("Dad")
scores_boxplot_player("Drew")
scores_boxplot_player("Ichigo")
scores_boxplot_player("Jeff")
scores_boxplot_player("Jenny")
scores_boxplot_player("Megan")
scores_boxplot_player("Mom")
scores_boxplot_player("Zach")
```

## Bias Heatmap

Each column represents the scores on one person's round that week. Higher scores in a cell mean that the player tends to gain more points than other players on that creator's rounds.

Estimated Score Per Round = Average Score + Round Difficulty + Bias

```{r}
bias_table_heatmap()
```

## Scores on a Creator's Round

This plot displays the average score on each round by week. Hover over a point to see the creator for that round.

```{r}
creator_overlay_plot()
```

Each of the following plots shows the scores on the rounds created by one person each week. The red line represents the average for that week, and the blue dots are individual scores that week (they need to be labeled at some point):

```{r paged.print=TRUE}
generate_creator_plot("Alex")
generate_creator_plot("Chris")
generate_creator_plot("Dad")
generate_creator_plot("Drew")
generate_creator_plot("Ichigo")
generate_creator_plot("Jeff")
generate_creator_plot("Jenny")
generate_creator_plot("Megan")
generate_creator_plot("Mom")
generate_creator_plot("Zach")
```

## Scores by Player

Each of the following plots shows the final score for a single player each week. Notice the p-value in the bottom right side of each output summary. If this value is less than 0.05, this result is considered unlikely to happen by pure chance and may indicate an underlying trend. 

```{r paged.print=TRUE}
generate_player_plot("Alex")
generate_player_plot("Chris")
generate_player_plot("Dad")
generate_player_plot("Drew")
generate_player_plot("Ichigo")
generate_player_plot("Jeff")
generate_player_plot("Jenny")
generate_player_plot("Megan")
generate_player_plot("Mom")
generate_player_plot("Zach")
```